Exclusions

### 2022-01-29 exclusions.Rmd
fixed up discussion

---
title: "Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon"
output:
  bookdown::html_document2:
   # reference_docx: wordstyles.docx
    toc: true
    toc_float: true
    number_sections: true
    css: custom.css
    code_folding: hide
always_allow_html: true
bibliography: sources.bib

@@

```{r packages, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(report)
library(lmerTest)
library(plotly)
library(corrr)
library(boot)
library(zeallot)
library(pander)
library(sjPlot)
library(ggeffects)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)
```

**Abstract**

Iconicity in language is receiving increased attention from many fields, but our understanding of the roles of iconicity in language is only as good as the measures we use to quantify it. We conducted iconicity rating and guessing experiments with 304 Japanese ideophones and prosaic words with sensory meanings, e.g. *fuwafuwa* ‘fluffy’, *jawarakai* ‘soft’. For both word groups, ratings and guesses were positively correlated—suggesting the two measures pick up on the same associations. Ideophones were consistently associated with higher iconicity ratings, but not higher guessing accuracy. We suggest that the structural markedness of ideophones enhances their perceived iconicity in the rating task, but does not provide any advantage (over and above form-meaning associations) in the guessing task. Thus, guesses and ratings could be used together to tease apart the relative contribution of structural markedness to iconic effects. Some ideophones were also poorly guessed, highlighting that construals of iconicity can be language-specific. Finally, we present some methodological contributions, including a new guessing paradigm that improves on the robustness, sensitivity and discriminability of previous approaches, and a reproducible workflow for creating rating and guessing experiments with a Python package, `icotools`, which we hope will improve comparability between future studies.

**Keywords:** iconicity, ideophones, lexical norms, iconicity ratings, guessability

# Introduction

What does it mean when we say that something is a "good word for" a concept? Certain words, like *klutz*, *schmooze*, and *smush*, seem to "suit" their meanings. We may have a sense that the French *belle* is a more appropriate word for BEAUTY than the Swedish *vacker*. Such intuitions are likely to be somewhat subjective, and influenced by our own linguistic experiences, but we all have them. The idea that certain forms can be particularly suited to certain meanings is known as *iconicity*.

Iconicity--defined here as a perceived resemblance between aspects of form and meaning--is now increasingly recognised as a *key property of language* [see e.g. @perniss_iconicity_2010; @perniss_bridge_2014; @dingemanse_arbitrariness_2015 for a review], raising questions about where iconic mappings come from, and what they do. To answer these questions, we first need to be able to *identify* iconicity in language and, ideally, to *measure* it.

Approaches to measuring iconicity can be divided into descriptive, data-driven, and behavioural methods [@motamedi_iconicity_2019]. Of these, only descriptive approaches target *iconicity* directly, with other methods simply identifying form-meaning association *biases*--whether these be based on resemblance (as in iconicity), co-occurrence (indexicality), statistically driven associations (systematicity), functional advantages, or other factors.

With descriptive approaches, iconicity as a theoretical concept is coded for directly (e.g. @pietrandrea_iconicity_2002; @hwang_body_2017; @ostling_visual_2018; @thompson_defining_2019; @voronin_osnovy_2006; @flaksman2020stanislav). This is more common and straightforward for sign languages than for spoken languages [@thompson_defining_2019]. It is usually done manually, but Östling et al. -@ostling_visual_2018 have shown for sign languages that (for some mappings at least) this can also be automated. Such studies usually involve cross-linguistic data, and aim to uncover universalities in how different languages map sound to meaning--that is, *where* iconicity comes from.

A wider variety of approaches target iconicity *indirectly* through related phenomena. For example, data-driven approaches (e.g. @wichmann_sound_2010; @blasi_soundmeaning_2016; @joo_phonosemantic_2019; @johansson_typology_2020; @winter_trilledr) use comparisons across unrelated languages to identify form-meaning correspondences that occur at rates higher than chance. This requires large amounts of parallel data from distinct phyla, and existing studies usually focus on basic vocabulary--where such data is most readily available. Like descriptive approaches, data-driven approaches identify *specific* form-meaning associations involved in iconic effects, but with the added advantage of not needing to rely on the intuitions and manual labour of individual researchers. However, they *do* rely on the existence of translation equivalents in multiple languages for the words under study. This can pose a particular challenge for highly iconic words, as these often also have highly specific meanings (@lupyan_language_2018; @akita_toward_2012). Thus, these approaches are less likely to be helpful for those with *specific* words for which iconicity measures are required.

Another body of work which capitalises on big data involves the study of *affective iconicity* [e.g. @ullrich_phonological_2016; @aryani_why_2018]. These studies use large datasets of affective norms to identify phonological segments loaded with affective meaning (by taking average affectiveness values across all words containing those segments, and identifying segments with values deviating from the global neutral mean). Although limited to associations involving affect, this approach has been successfully used to investigate iconic effects in language processing [@schmidtke_effects_2018; Ullrich in prep; @aryani_affective_2018], and has even been used to explain the affective meaning of poems [@ullrich_relation_2017].

Behavioural approaches involve the collection of data from naive experimental participants, which can reveal something about the iconicity of given form-meaning pairings. The only requirement is the human participants for the study, and the largest amount of iconicity measurements from the widest variety of languages have been collected using these methods. There are two main approaches: iconicity rating tasks and guessing experiments.

In iconicity rating tasks, the concept of iconicity is explained to participants, and participants are asked to rate the iconicity of given form-meaning pairs directly [e.g. @vinson_british_2008; @perry_iconicity_2015; @winter_which_2017; @occhino_iconicity_2017; @thompson_iconicity_2020; @punselie_iconicity_2020]. Measurements from these studies have been found to correlate in meaningful ways with factors related to iconicity--e.g. encoding of sensory information [@winter_which_2017], semantic domains [@punselie_iconicity_2020], semantic neighbourhood density [@sidhu_lonely_2018], word class [@perry_iconicity_2015; @thompson_iconicity_2020], age of acquisition [@vinson_british_2008; @perry_iconicity_2015], facilitated processing [@thompson_link_2009; @occhino_role_2020], funniness and structural markedness [@dingemanse_playful_2020]. However, ratings given are influenced by the linguistic experience of participants (@occhino_iconicity_2017; @sevcikova_sehyr_perceived_2019), so some care is needed to ensure measures are taken from appropriate participant groups. Further, ratings from native speakers may also capture non-iconic regularities like semantic transparency and indexicality [@dingemanse_playful_2020; @winter_iconicity_2021], so cannot be taken as direct reflections of degree of iconicity.

In guessing experiments, participants are asked to guess the meaning of foreign/novel signs from a choice of two or more alternatives [e.g. @kohler_gestalt_1929; @tsuru_problem_1933; @brown_phonetic_1955; @kunihira_effects_1971; @ramachandran_synaesthesia_2001; @dingemanse_what_2016; @tzeng_specificity_2017; @perlman_iconicity_2015; @perlman_people_2018; @lockwood_sound-symbolism_2016]. Signs whose meanings are correctly guessed at rates higher than chance are considered iconic. The guessability of signs has also been linked to facilitated processing [@ormel_role_2009] and learnability [@perlman_people_2018; @lockwood_sound-symbolism_2016], and correlated with iconicity ratings [@punselie_iconicity_2020]. However, some signs that were given low iconicity ratings were actually highly guessable, while there were also a few signs that, despite receiving high ratings, were not easily guessed  [@punselie_iconicity_2020: 23]. This suggests that guessing and rating tasks may tap into slightly different constructs, or be sensitive to experimental design constraints.

While results from behavioural experiments often require some untangling, we believe that they are particularly valuable sources of iconicity measures for four reasons: (1) they can be applied to any kind of dataset, (2) they are scalable, being particularly amenable to online, crowd-sourced data collection, (3) they are tweakable, meaning that we can experiment with different experimental designs to figure out exactly what these measures are telling us, and (4) they are amenable to standardisation, making them a good option for reproducibility. In fact, they are the *only* measures of iconicity that we know of to combine all these things.  Moreover, that the measures correlate with each other and with a variety of other meaningful phenomena speaks to their validity.

In this study, we seek to further our understanding of these measures and how they can be useful by comparing and contrasting iconicity ratings and guesses for 304 Japanese words from the sensory vocabulary of Japanese. Sensory language was chosen as this is a domain high in iconicity, and where iconicity offers clear advantages [@winter_which_2017]. The data contained a mix of ideophones--depictive, eye-catching words like *fuwafuwa* 'soft' and *pikapika* 'sparkly' that "invite and afford the construal of iconic mappings between form and meaning" [@akita_ideophone_2019: 18]--along with prosaic lexical items, like *yawai* 'soft' and *hikaru* 'shine', which are not necessarily expected to be lacking in iconicity--especially given their sensory meanings--but which are less obvious about it. We were particularly interested in how the iconicity measures for these words would compare between the rating studies--where participants are asked to make a conscious decision about the iconicity of the words--and the guessing studies--which tap into more unconscious biases. We find that when used together, the two methods shed light on each other, as well as raising pertinent questions about where iconicity effects come from. Our conclusion is that, for many of the questions that we have about iconicity in language, two measures may be better than one.

To discover how to get the most out of each measure, we compare and contrast different rating and guessing paradigms, and in particular introduce a new guessing paradigm that improves on the robustness, discriminability and sensitivity of previous approaches. Finally, to promote the use of these new measures, we introduce a reproducible workflow for automatically generating rating and guessing experiments in a standardised format, via a python package `icotools`, with support for a variety of stimulus formats (audio, video, and images). This will make the collection of behavioural based iconicity measures quicker, easier, and more comparable between future studies.

# Methods

The data from the study, as well as the code used to produce the experiments and analyses can be found on the Open Science Framework, at https://osf.io/j57uc/?view_only=d59ffde4bc42467fafb71a904bef8d14.

## Stimuli

The stimuli were 304 words taken from the *Japonic Sensory Lexicon* database (ANONYMISED CITATION), which is a collection of words for 110 sensory concepts in 49 Japonic varities.

```{r}
# semantic domains of concepts, plus numbers of ideophones and non-ideophones
concepts <- read_tsv("https://raw.githubusercontent.com/BonnieMcLean/JaponicSensoryLex/main/concepts.tsv")%>%
  rowwise()%>%
  mutate(concept=str_split(concept,"-")[[1]][1])%>%
  select(concept,domain)

words <- read_csv("responses-guessingwords.csv")%>%
  select(form,identifier)%>%
  mutate(identifier=str_replace_all(identifier,"FUZZY","CLEAR"))%>%
  rowwise()%>%
  mutate(concept=str_split(identifier,"_")[[1]][2])%>%
  unique()%>%
  left_join(concepts,by="concept")

words <- read_csv("ideophone_list.csv")%>%
  select(form,ideophone)%>%
  right_join(words,by="form")

# plus the practise and control items which arent in the Japonic sensory lexicon
words$domain[words$form == "yoroyoro"] <- "action"
words$domain[words$form == "fuwafuwa"] <- "texture"
words$domain[words$form == "hisohiso"] <- "sound"
words$domain[words$form == "pyoNpyoN"] <- "action"

```

```{r}
dat <- read_csv("https://raw.githubusercontent.com/BonnieMcLean/JaponicSensoryLex/main/dat.csv")
info <- 
words <- read_csv("responses-guessingwords.csv")%>%
  select(identifier)%>%
  mutate(identifier=str_replace_all(identifier,"FUZZY","CLEAR"))%>%
  unique()

cognates <- read_csv("cognatecodes.csv")%>%
  mutate(identifier=paste(test_form,eng_trans,sep="_"))%>%select(cognate,identifier)

words <- words%>%left_join(cognates,by="identifier")
```

The stimuli were 304 words from the Yamato stratum of the Japanese lexicon, in the semantic domain of sensory perception. The words were taken from the *Japonic Sensory Lexicon* (ANONYMISED CITATION), a dataset containing words in

from the domain of sensory perception, including both ideophones and prosaic lexical items. For all tasks, the words were presented to participants as audio files. The audio files were synthesised using the Google Cloud Text-to-Speech API [(https://cloud.google.com/text-to-speech)](https://cloud.google.com/text-to-speech), with the female Japanese Wavenet voice ja-JP-Wavenet-B. They were then edited using Praat [@boersma_praat_2020] to have a flat pitch of 200Hz. Expressive prosody can enhance performance in guessing tasks [@kunihira_effects_1971; @dingemanse_what_2016], and so using a flat intonation was seen as the best way to ensure comparisons between the words were fair and not influenced by the prosody in a particular recording. We also considered synthesising all the words to have a marked intonation. However, different word lengths would make changes in pitch more dramatic in shorter words compared to longer words, affecting words of different lengths unequally. While flat, synthetic voices may be the best option for comparability, they are not a natural reflection of real-world experiences with iconicity--which involve not only segments but also prosody and even multimodal cues (e.g. accompanying gesture). Thus, if anything the results in this and similar studies likely *underestimate* iconicity in language.

In all studies, participants were instructed to use headphones throughout the experiment, and could not procede to the experiment without first passing a listening test designed to require the use of headphones @woods_headphone_2017. The response buttons in each trial also did not appear until *after* the audio files had finished playing. This ensured that participants could not respond without first listening to the audio files.

## Data collection

All studies were conducted online, with participants recruited through the crowd sourcing service Prolific (www.prolific.co). The participants were monolingual English speakers residing in England, with no prior knowledge of Japanese. Although rating tasks are often conducted with native speakers, we used English speakers in this study for two reasons: first, to make the results as comparable as possible between the guessing and rating tasks; and second, because our data contained a mix of Standard Japanese and non-standard Japanese words, and we were concerned that using Standard Japanese speakers would lead to a bias towards standard words [@occhino_iconicity_2017].

The words were split across 11 different experiments, in such a way that each experiment had a balance of words that were expected to be iconic, versus words that were not expected to be iconic. In addition, two 'practice words' and three 'control words' were included in every experiment. The practice items were used to give participants time to familiarise themselves with the experiment design, and were not included in the results. The control items were chosen from words found to be very iconic/highly guessable in previous studies, and were included as a check to identify participants who were perhaps not performing the task as intended. The first two trials in every experiment were the two 'practice words', while the order of the rest of the words in the experiment, including the control words, was randomly varied between participants. After finishing the experiment, participants were asked to describe the task they were performing during the experiment. This was used as an attention check, and to identify participants who had misunderstood the experiment instructions. Strange or lacking task descriptions, particularly when coupled with poor performance on the control items, were used to identify and exclude results from participants who were likely not performing the task as intended.

@@

For the rating task, data from 5 out of 225 participants was excluded. For the ratings, as well as looking at task descriptions and performance on the control items, we calculated the person-total correlation for each participant @curran_methods_2016. The person-total correlation is the correlation between an individual participant's ratings, and the mean ratings from every other participant. It provides a measure of how well a participant's responses agree with those of other participants in the task. The criteria for exclusion was two or more of either (a) strange or lacking task descriptions, (b) very low ratings for control items, or (c) low person-total correlations (<.2^[We chose .2 as the cut off point because it corresponded with a discontinuity in the lower end of the distribution of person-total correlations, see Figure 1 in the online Supplementary Material]). Criteria (a) was used for 4 exclusions. In two cases, the participants indicated that they were rating the words according to how much they sounded like their English translations, rather than how well they resembled their meanings. Criteria (b) was used in 3 exclusions, and criteria (c) was used in 4 exclusions.

```{r preprocessing ratings}
# for ratings
ratings <- read_csv("responses-ratings.csv")

@@

persontotal_dist <- ggplot(data=persontotals,aes(x=pt_corr))+geom_histogram()+scale_x_continuous(n.breaks=8)

@@

Each experiment took a median of 6 minutes to complete, and participants were paid £0.75 for their participation (based on Prolific's 'good' hourly rate of £7.50). Prolific's prescreening features were used to ensure that participants who had already participated in one experiment (regardless of whether it was a rating or guessing task) could not participate in any subsequent experiments. Excluding the practice and control items (which every participant guessed and rated), in the final dataset we had a median of 40 guesses per word (range: 15-129) in the pilot guessing experiment; a median of 30 guesses (range: 29-60) per word for the second guessing experiment; and a median of 15 ratings (range: 14-58) per word for the rating task (@motamedi_iconicity_2019 suggest a benchmark of 10 ratings per word is sufficient).

```{r final data count}
ratings%>%
  select(identifier,prolificID)%>%
  unique()%>%
  group_by(identifier)%>%
  count()%>%
  filter(n<200)->no_ratings

#median(no_ratings$n)

guesses_words%>%
  select(identifier,prolificID)%>%
  unique()%>%
  group_by(identifier)%>%
  count()%>%
  filter(n<200)->guess1

# median(guess1$n)

guesses_trans%>%
  mutate(identifier=paste(word,concept))%>%
  select(identifier,Prolific.ID)%>%
  unique()%>%
  group_by(identifier)%>%
  count()%>%
  filter(n<200)->guess2

#median(guess2$n)  
```

## Pilot guessing experiments: guessing between translations

We piloted some guessing experiments following previous studies [e.g. @dingemanse_what_2016], in which participants were presented with a Japanese word and asked to guess its meaning from a choice of two possible translations. Some of the ideophones in the data had very specific meanings, which were difficult to capture in a single English word. To provide more context, the translations were given in sentence frames, with the translation word in capitals. The instructions were as follows:

> We are interested in how well people can guess the meanings of words in foreign languages. In this experiment, we are interested in your intuitions about the meanings of JAPANESE words.
You will hear 25 Japanese words. After listening to each word, try to guess its meaning from a choice of two possible English translations. For context, meanings are given in a sentence, with the intended meaning in CAPITALS.
For example, if you think the word could mean BRIGHT in the sentence 'The sun is BRIGHT', choose that sentence. If you think it could mean ROUND in the sentence, 'The sun is ROUND', choose that sentence. Trust your gut, and good luck!

A sample trial is shown in Figure \@ref(fig:guessingtrans).

```{r guessingtrans, fig.cap="Guessing task - guessing between translations",echo=FALSE}
knitr::include_graphics("images/guessingtransdemo.PNG")
```

The same formulaic structure, "After X, Y", was used for all the sentences--where Y is the target perception and X is the event that causes it. As some sensory modalities (e.g. sound) may better lend themselves to the construal of iconicity than others (e.g. taste), correct translations were always paired with foil translations from the same modality (e.g. sound translations were only paired with other sound translations, and taste translations with other taste translations, etc.). To make the difficulty of the task comparable between different trials, only words that were neither synonyms nor antonyms of the correct translation were chosen as foils. Finally, the length of the sentences was also kept consistent (to a difference of no more than 5 characters) between pairings of translations and foils. 
To test the robustness of the guessing procedure to different choices of translation and foil words, the same words were tested multiple times with different translations and foils. We were concerned that participants could choose particular translations because of how they sound (particularly if they sounded similar to the Japanese word), rather than based on their meanings as we intended. So, to minimise any potential effects from this, as much as possible phonologically distinct synonyms were chosen as alternative translations (e.g. SMALL|TINY, BIG|LARGE).

Pilot results showed that in several cases the guessability of a word did differ considerably depending on the translation or foil word used. Some examples are shown in Table \@ref(tab:weirdcases).

```{r weirdcases,echo=FALSE}
guesses_trans%>%
  mutate(word_spec=paste(word,concept,sep="_"))->guesses_trans

guesses_trans%>%
  filter(word_spec=="piiN_LONG"|word_spec=="hakkiri_CLEAR-HEADED")%>%
  filter(!(grepl("SHARP",answer_sen)))%>%
  select(answer_sen,foil_sen,answer)%>%
  group_by(answer_sen,foil_sen)%>%
  count(answer)%>%
  pivot_wider(names_from=answer,values_from=n, values_fill=0)%>%
  knitr::kable(col.names = c("Answer sentence", "Foil sentence","correct","incorrect"),caption="Effect of different foils and translations on guessing results for hakkiri 'CLEAR HEADED' and piiN 'LONG'")%>%
  kable_styling()

```

We hypothesised that, as well as considering the meaning of the translations and foils, participants may be driven to one option over another because of the way the words sound (particularly if they sounded similar to the Japanese word being tested), or because of lexical features of the words like frequency, valence, arousal, or any other associations they may have with these words.

While we could not do anything about the sound of the words (other than using multiple translations), we realised that we could eliminate at least some of the complications caused by other properties of the words if we flipped the design of the guessing experiments on its head, so that instead of choosing the correct English meaning for a given Japanese word, participants chose the correct Japanese word for a given English meaning. In this way, there was only one English word involved in the task, and the options participants chose between were simply two Japanese words which they knew nothing about, other than how they sound. This design was also more comparable to the rating task--where participants are also given the meaning of the word first. We therefore decided to use this new design for the rest of the guessing experiments.

## Guessing experiments: guessing between words {#goodguesses}

For the new version of the guessing experiments, we minimised the amount of English words involved in the task by having participants match a single English translation to the corresponding word in Japanese, choosing between two possible Japanese words. The instructions used were as follows:

> We are interested in how well people can guess words in foreign languages. In this experiment, you will be asked to match the English translation with the corresponding word in JAPANESE, guessing from a choice of two Japanese words.

A sample trial is shown in Figure \@ref(fig:guessingwords).

```{r guessingwords, fig.cap="Guessing task - guessing between words",echo=FALSE}
knitr::include_graphics("images/guessingwords.PNG")
```

This time, instead of using sentence frames for context, we kept the amount of English to a minimum by using only single-word translations where possible, and where more context was needed this was provided in brackets, as in Figure \@ref(fig:guessingwordsspec).

```{r guessingwordsspec, fig.cap="Guessing task - guessing between words",echo=FALSE}
knitr::include_graphics("images/guessingwordsspec.PNG")
```

The incorrect 'foil' word in each item was a nonsense Japanese word artificially generated to be as phonologically distinct as possible from the correct word, while still conforming to Japanese phonology. For every word, three different foil words were generated by substituting each consonant with its top three most phonologically distant consonants in the Japanese sound system. Phonological distances were calculated using the feature matrix provided in PHOIBLE [@phoible], with the distance between two sounds being equal to the sum of the distances between each of their feature values. Distances were calculated as follows:

> The distance between two feature values that are identical is 0. The distance between two opposing values (+/- or -/+) is 1—except if the feature involved is length or voicing, then the distance is 0.5, The distance between two feature values when one of them is 0 (= not applicable) is 0.25.

Since /a/ is often the most phonologically distant vowel from many of the vowels in the Japanese sound system, to avoid the foil words all having similar vowels, a different approach was used for substituting vowels. Instead of using feature values, vowels were substituted with the vowel obtained by rotating the vowel space 180 clockwise or counterclockwise. So /u/ is substituted with either /a/ or /i/, and /a/ is substituted with either /i/ or /u/ (see Figure \@ref(fig:vowels)). Finally, reduplicated words were ‘unreduplicated’ when creating foil words. For example, the foils for *fuwafuwa* were *watʃi*, *nati* and *ridːu* (not *watʃiwatʃi*, *natinati* and *ridːuridːu*).

```{r vowels,fig.cap="Vowel substitutions for the creation of foil words",echo=FALSE}
knitr::include_graphics("images/vowels.PNG")
```

By using foil words that sound as different as possible to the target word, we hoped to improve the sensitivity of the measure to any iconicity in the target word. Our logic was that if the target word *is* iconic for a concept, then a word that sounds very different to it should be a bad fit for that same concept--making the choice between the two easier. To test whether this was really the case, we ran one of the experiments a second time using foils that were randomly chosen from among the other trial items in the experiment, rather than phonologically distinct foils.

## Rating task {#ratings}

We also conducted an iconicity rating task with the same words. In the rating task, the concept of iconicity was defined to participants as "when a word and its meaning resemble one another", using the English examples *wiggle*, *jiggle*, and *wriggle*. The instructions stated that "Even people who do not speak any English can get a sense of the meaning of these words", and contrasted them with words like *walk* and *run* whose meanings are "not so intuitive". The full instructions given were as follows:

> Some words seem to 'fit' their meanings. For example, consider the English words wiggle, jiggle, and wriggle.  
We have an intuitive sense of the meanings of these words, because there is a resemblance between the words and their meanings.  
Even people who do not speak any English can get a sense of the meaning of these words.  
Words like walk and run on the other hand are not so intuitive; people who do not know any English would not be able to guess what these words mean.  
In this task, you will listen to some Japanese words, and we will tell you their meanings. You will then be asked to judge whether there is a resemblance between the word and its meaning.

Participants had to listen to the Japanese words and were told their meanings, then asked to rate the resemblance between the word and its meaning on a scale from 0 'No resemblance' to 6 'Strong resemblance' (see Figure \@ref(fig:ratingdemo)). Previous studies [@perry_iconicity_2015] also used a negative end of the scale (corresponding to a 'bad resemblance'). However, follow-up analyses found that the negative end of the scale was both underused and inconsistently used, suggesting that anti-iconic relationships are difficult to assess [@motamedi_iconicity_2019]. For this reason, we decided to only use a positive scale in this study.

```{r ratingdemo, fig.cap="Rating task",echo=FALSE}
knitr::include_graphics("images/ratingdemo.PNG")
```

As with the guessing experiments, the translation shown for each participant was varied randomly between a set of (wherever possible) phonologically distinct synonyms.

## Creation of Icotools

To make the collection of guesses and ratings quicker and easier for future studies, while also improving comparability between studies, we created a python package, `icotools` [(https://pypi.org/project/icotools/)](https://pypi.org/project/icotools/), with functions that can be used to automatically generate rating and guessing experiments (using the meaning-to-word design from the second guessing experiments). The package has three functions: a rater function, a guesser function, and a foiler function. The rater and guesser functions create rating and guessing experiments as described in Sections \@ref(goodguesses) and \@ref(ratings). They take two csv files as input. The first is a wordlist containing the items in the experiment (including trials, practice items, control items, and sets of translations to use for them), while the second file is used for customisation (e.g. to specify how many items to include in each experiment, the location of media files, custom instructions and custom exit questions, etc.).

For each experiment, the functions produce three files--a html file, a php file, and a csv file. These files can then be uploaded to a server, the link to the html files can be shared with participants, and as the participants complete the experiments their responses will be written to the csv file by the php file. The functions also produce a summary csv file with all the details of the experiments (i.e. the trials in each experiment, as well as the translations and, where applicable, foils used). An R script is provided to collate the responses for all experiments into one long tidy dataset after data collection is complete.

The foiler function can be used to generate a list of opposite sounding foil words for each item in a guessing experiment, following the method described in Section \@ref(goodguesses). It is also possible to specify your own foil words, or to choose the foils randomly from among the other items in the experiment. For a full description of how to use icotools, please see the ReadMe at https://anonymous.4open.science/r/IcoTools-2CEA/README.md (this is an anonymised link to the github repository: it will later be replaced with the correct link).

# Results

Statistical analyses were conducted using `r R.Version()$version.string`.

## Robustness to different choices of translations and foils

We first investigated whether ratings and guesses were robust to different choices of translations (for ratings and guesses) and foils (for guesses only). For each word separately, we performed Fisher's Exact Tests on the guessing data, and ANOVAs on the rating data, to determine whether the guessing performance or ratings given differed significantly between different translations and foils. As the ANOVAs are less robust to small amounts of data than the Fisher's Exact Tests, for the rating data we excluded translations with fewer than 10 ratings from this particular analysis.

```{r test robustness guessing experiments}
# for guessing between translations
words<-unique(guesses_trans$word_spec)

# see whether the responses differed significantly depending on the answer sentence given
n=0
doubly_tested_words <- c()
weird_words<-c()
for(w in words){
  d<-subset(guesses_trans,word_spec==w)
  t<-table(d$answer_sen,d$answer)
  # not all words were tested with more than one translation as for some concepts there was only one English word for the concept
  if (nrow(t)>1){
    doubly_tested_words <- c(doubly_tested_words,w)
    n=n+1
    p_val<-fisher.test(t)$p
    if (p_val<0.05){
      weird_words<-c(weird_words,w)}
}
}

# see whether the responses different signficantly depending on the foil sentence given
for(w in words){
  d<-subset(guesses_trans,word_spec==w)
  t<-table(d$foil_sen,d$answer)
  if (nrow(t)>1){
    doubly_tested_words <- c(doubly_tested_words,w)
    n=n+1
    p_val<-fisher.test(t)$p
    if (p_val<0.05){
      weird_words<-c(weird_words,w)}
}
}

weird_res_trans <- length(unique(weird_words))/length(unique(doubly_tested_words))

# for guessing between words

words<-unique(guesses_words$identifier)

# see whether the responses differed significantly depending on the answer sentence given
n=0
doubly_tested_words <- c()
weird_words<-c()
for(w in words){
  d<-subset(guesses_words,identifier==w)
  t<-table(d$trans,d$answer)
  if (nrow(t)>1&ncol(t)>1){
    doubly_tested_words <- c(doubly_tested_words,w)
    n=n+1
    p_val<-fisher.test(t)$p
    if (p_val<0.05){
      weird_words<-c(weird_words,w)}
}
}

# see whether the responses different signficantly depending on the foil sentence given
for(w in words){
  d<-subset(guesses_words,identifier==w)
  t<-table(d$foil,d$answer)
  if (nrow(t)>1 & ncol(t)>1){
    doubly_tested_words <- c(doubly_tested_words,w)
    n=n+1
    p_val<-fisher.test(t)$p
    if (p_val<0.05){
      weird_words<-c(weird_words,w)}
}
}

weird_res_words <- length(unique(weird_words))/length(unique(doubly_tested_words))
```

```{r ratings robustness}
# for the ratings we are using anovas which are not as robust to small amounts of data as the fisher test

# therefore I only looked at data where we had at least ten ratings per translation
ratings%>%
  select(identifier,trans,rating)%>%
  group_by(identifier)%>%
  add_count(trans)%>%
  filter(n>=10)->data

# then within this, we need to get the data where we have at least two translations per word
data%>%
  select(identifier,trans)%>%
  unique()%>%
  group_by(identifier)%>%
  mutate(count=n())%>%
  arrange(identifier)%>%
  filter(count>1)%>%
  pull(identifier) -> test

data <- subset(data,identifier %in% test)

words <- unique(data$identifier)
weird_res <- c()
for(w in words){
  d <- subset(data,identifier==w)
  n_trans <- length(unique(d$trans))

  if(n_trans==2){
  t <- t.test(d$rating~d$trans)
  p <- t$p.value
  }
  else{
    t <- summary(aov(d$rating~d$trans))
    p <- t[[1]][[5]][1]
  }

  if(p<0.05){weird_res <- c(weird_res,w)}
}

weird_res_ratings <- length(weird_res)/length(words)
```

In the first pilot of the guessing task--where participants were guessing between English translations--11 out of the 45 words tested (=24%) showed significant differences in their guessability when either the translation and/or the foil was changed. In the second version of the guessing task--where participants were guessing between Japanese words--39 out of 303 words tested (=13%) showed significant differences in their guessability when either the translation and/or the foil was changed. In the rating task, 11 out of 88 words tested (=13%) showed significant differences in their iconicity rating when the translation word was changed.

## Sensitivity and discriminability

To test whether using phonologically distinct foils in the guessing experiments improved their sensitivity to iconicity, a subset of 30 words were tested twice--once with phonologically distinct "opposite" foils, and once with random foils.

The number of these words which were guessed significantly above chance (with a sample of 30 guesses per word, an accuracy greater than 2/3 is needed to be sure the true accuracy is above 0.5) was higher in the experiment with opposite foils than in the experiment with random foils. Using opposite foils, half of the words in the sample (15/30) were guessed significantly above chance, whereas with the random foils this number fell to just 1 in 5 words (6/30). Only two words were guessed significantly above chance in the random foil condition, but not in the opposite foil condition. However, both were guessed numerically above chance in the opposite foil condition as well (with accuracies of 52% and 57%). The differences between the two conditions are shown in Figure \@ref(fig:sensitivitycheck). A chi-squared test confirmed the differences to be significant (χ2(1)=4.69,p=.03).

```{r sensitivitycheck,fig.cap="Words guessed significantly above chance using opposite versus random foils"}

# we used this binomial test to figure out how many correct guesses (out of 30) were needed to have a result significantly different from chance--it's 21
#binom.test(21,30,p=.5)
randomfoils <- read_csv("responses-randomfoils.csv")

randomfoils%>%
  group_by(identifier)%>%
  mutate(score=sum(result=="correct")/sum(result=="correct"|result=="incorrect"))%>%
  select(form,score)%>%unique()->randomres

# fix up some naming differences
words <- randomres$identifier
words <- replace(words,words=="zoNzoN_SHIVERING","zoNzoN_(SPINE) TINGLING")
words <- replace(words,words=="geNki_LIVELY","geNki_LIVELY (ENERGETIC)")

guesses_words%>%
  filter(identifier %in% words|form=="zyiNzyiN")%>%
  group_by(identifier)%>%
  mutate(score=sum(answer=="correct")/sum(answer=="correct"|answer=="incorrect"))%>%
  ungroup()%>%
  select(form,score)%>%unique()->oppres

comparison <- left_join(randomres%>%select(-identifier),oppres,by="form")%>%
  rename(random_foils=score.x,opp_foils=score.y)%>%unique()

comparison%>%
  rename(`random foils`=random_foils,`opposite foils`=opp_foils)%>%
  pivot_longer(c("random foils","opposite foils"),names_to="design",values_to="accuracy")%>%
  mutate(`guessing accuracy`=ifelse(accuracy>2/3,"significantly above chance","at chance"))%>%
  ungroup()%>%
  select(design,`guessing accuracy`)%>%
  group_by(design)%>%
  count(`guessing accuracy`)%>%
  ggplot(aes(x=design,y=n,fill=`guessing accuracy`))+geom_col()+theme_classic()+labs(x="Design",y="Number of words")+scale_fill_manual(values=c("#CCCCCC","#000000"))
```

```{r chi square}
comparison%>%
  rename(`random foils`=random_foils,`opposite foils`=opp_foils)%>%
  pivot_longer(c("random foils","opposite foils"),names_to="design",values_to="accuracy")%>%
  mutate(`above chance`=ifelse(accuracy>2/3,"y","n"))->comparison

#table(comparison$design,comparison$`above chance`)
#chisq.test(table(comparison$design,comparison$`above chance`))
```

>>>>

---
title: "Exclusions"
output: 
    bookdown::html_document2:
      toc: true
      toc_float: true
      number_sections: true
      css: custom.css
      code_folding: hide
bibliography: sources.bib

@@

```{r packages, include=FALSE}
library(bookdown)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)
```

@@

For the rating task, data from 5 out of 225 participants was excluded. For the ratings, as well as looking at task descriptions and performance on the control items, we calculated the person-total correlation for each participant [@curran_methods_2016]. The person-total correlation is the correlation between an individual participant's ratings, and the mean ratings from every other participant. It provides a measure of how well a participant's responses agree with those of other participants in the task. The criteria for exclusion was two or more of either (a) strange or lacking task descriptions, (b) very low ratings for control items, or (c) low person-total correlations. Criteria (a) was used for 4 exclusions. In two cases, the participants indicated that they were rating the words according to how much they sounded like their English translations, rather than how well they resembled their meanings. Criteria (b) was used in 3 exclusions, and criteria (c) was used in 4 exclusions. A person-total correlation below .2 was considered a 'low' correlation, based on the distribution of person-total correlations shown in Figure \@ref(fig:ratingsprocess).

```{r ratingsprocess,fig.cap="Distribution of person-total correlations in rating task"}
# for ratings
ratings <- read_csv("responses-ratings.csv")

@@

persontotal_dist <- ggplot(data=persontotals,aes(x=pt_corr))+geom_histogram()+scale_x_continuous(n.breaks=8)+labs(x="Person-total correlation")
persontotal_dist

@@
